SMT-Boosted Security Types and Correctness for Low-Level MPC

### 2025-11-24 toplas26/intro.tex
checkpoint

Data privacy is a critical concern in distributed applications
including privacy-preserving machine learning
\cite{li2021privacy,knott2021crypten,koch2020privacy,liu2020privacy}
and blockchains
\cite{ishai2009zero,lu2019honeybadgermpc,gao2022symmeproof,tomaz2020preserving}.
Secure multiparty computation (MPC) is a key software-based enabling
technology for these applications. MPC protocols provide both confidentiality
and integrity properties, formulated as real/ideal (aka simulator)
security and universal composability (UC) \cite{evans2018pragmatic},
with well-studied manual proof methods \cite{Lindell2017}.  MPC
security semantics has also been reformulated as
\emph{hyperproperties}
\cite{8429300,10.1145/3453483.3454074,skalka-near-ppdp24} to support
automated language-based proof methods. Recent research in the SMT
community has developed theories of finite fields \cite{SMFF} with
clear relevance to verification of cryptographic schemes that rely on
field arithmetic. The goal of this paper is to combine these two
approaches in a decidable type system for verifying correctness and
security properties of MPC protocols.

@@

Our work is based on the $\metaprot$/$\minicat$ framework developed in
previous work \cite{skalka-near-ppdp24, skalka-near-esop25}. The prior work developed the 
language and semantics, plus two mechanisms for verifying security: an
automated approach for protocols in $\mathbb{F}_2$ (the binary field),
based on enumerating all possible executions of the protocol, that scales 
only to very small protocols; and a manual approach based on a program logic.
%
In Section \ref{section-lang} we recall the $\minicat$ language model and key definitions. In
Section \ref{section-model} we recall formulations of program distributions,
and key probabilistic hyperproperties of MPC in $\minicat$.

>>>>

Data privacy is a critical concern in distributed applications
including privacy-preserving machine learning
\cite{li2021privacy,knott2021crypten,koch2020privacy,liu2020privacy}
and blockchains
\cite{ishai2009zero,lu2019honeybadgermpc,gao2022symmeproof,tomaz2020preserving}.
Secure multiparty computation (MPC) is a key software-based enabling
technology for these applications. MPC protocols provide both confidentiality
and integrity properties, formulated as real/ideal (aka simulator)
security and universal composability (UC) \cite{evans2018pragmatic},
with well-studied manual proof methods \cite{Lindell2017}.  MPC
security semantics has also been reformulated as
\emph{hyperproperties}
\cite{8429300,10.1145/3453483.3454074,skalka-near-ppdp24} to support
automated language-based proof methods. Recent research in the SMT
community has developed theories of finite fields \cite{SMFF} with
clear relevance to verification of cryptographic schemes that rely on
field arithmetic. The goal of this paper is to combine hyperproperty
formulations of security with SMT verification as a foundation for
enforcing correctness and security properties of MPC protocols.

@@

Our work is based on the $\metaprot$/$\minicat$ framework developed in
previous work \cite{skalka-near-ppdp24, skalka-near-esop25}. The prior work developed the 
language and semantics, plus two mechanisms for verifying security: an
automated approach for protocols in $\mathbb{F}_2$ (the binary field),
based on enumerating all possible executions of the protocol that scales 
only to very small protocols; and a manual approach based on a program logic.
%
In Section \ref{section-lang} we recall the $\minicat$ language model and key definitions. In
Section \ref{section-model} we recall formulations of program distributions,
and key probabilistic hyperproperties of MPC in $\minicat$.
